---
title: "Reactions to race analyses"
output:
  html_notebook: default
  html_document: default
---

#Reading and Tidying Data
```{r, message=FALSE, warning=FALSE}
#load pacakges
library(haven); library(tidyverse); library(forcats)
setwd("C:/Users/Daniel/Desktop/Konsulentarbeid/R2R/By_State")

#reads data sets by state
temp = list.files(pattern="*.sas7bdat")
for (i in 1:length(temp)) assign(temp[i], read_sas(temp[i]))

#make variable names consistent across data frames (states)
dfs <- mget(temp)

for (i in 1:length(dfs)) {
      cnames <- tolower(colnames(dfs[[i]]))
      colnames(dfs[[i]]) <- cnames
}

brfss.all <- bind_rows(dfs)

#define value labels
labs.age = c("Age 18 - 24", "Age 25 to 34", "Age 35 to 44",
             "Age 45 to 54", "Age 55 to 64", "Age 65 or older")

labs.bp = c("Yes","No")

labs.edu = c("Did not graduate High School", "Graduated High School",
             "Attended College or Technical School",
             "Graduated from College or Technical School")

labs.genhealth = c("Excellent", "Very good", "Good", "Fair", "Poor")

labs.health = c("Good", "Poor")

labs.hvydrink = c("Not at risk", "At risk")

labs.inc = c("Less than $15,000", "$15,000 to less than $25,000",
             "$25,000 to less than $35,000", "$35,000 to less than $50,000",
             "$50,000 or more")

labs.hlthdays = c("None", "1 - 13 days", "14 or more days")
	
labs.race2fmt = c("White only, non-Hispanic", "Black only, non-Hispanic",
                  "Asian only, non-Hispanic", 
                  "Native Hawaiian or other Pacific Islander only, No",
                  "American Indian or Alaskan Native only, Non-Hispan",
                  "Other race only, non-Hispanic",
                  "Multiracial, non-Hispanic",
                  "Hispanic", "Don't know/Not sure/Refused")

labs.region = c("South", "Northeast", "Midwest", "Southwest", "West")

labs.rr2atwrk = c("Worse than other races", "The same as other races",
                  "Better than other races", 
                  "Worse than some races, better than others",
                  "Only encountered people of the same race")

labs.rr2cognt = c("Never", "Once a year", "Once a month", "Once a week",
                  "Once a day", "Once an hour", "Constantly")

labs.rr2yesno = c("Yes", "No", "Don't know/Not Sure", "Refused")
labs.rr2yesno2 =  c("Yes", "No")

labs.rr3hcare = c("Worse than other races", "The same as other races",
                  "Better than other races", 
                  "Worse than some races, better than others",
                  "Only encountered people of the same race",
                  "No health care in past 12 months") 

labs.rrclass = c("White", "Black or African American", "Hispanic or Latino",
                 "Asian", "Native Hawaiian or Other Pacific Islander",
                 "American Indian or Alaska Native", "Some Other Group")

labs.rrclass2 = c("White", "Black or African American", "Hispanic or Latino",
                 "Asian", "Native Hawaiian or Other Pacific Islander",
                 "American Indian or Alaska Native")

labs.sex = c("Male", "Female")

labs.smoker = c("Current smoker", "Former smoker", "Never smoked") 
labs.smoker2 = c("Current smoker", "Former smoker", "Never smoked", "DK") 

##Assign values labels to values across all data frames (states)
brfss.all <- brfss.all %>% mutate(age_gr=factor(`_age_g`, labels=labs.age),
                                gen_health=ordered(genhlth, labels=labs.genhealth),
                                educ=factor(`_educag`, labels=labs.edu),
                                hvy_drink=factor(hvydrink, labels=labs.hvydrink),
                                income=factor(`_incomg`, labels=labs.inc),
                                ment_days=factor(mentdays, labels=labs.hlthdays),
                                phys_days=factor(physdays, labels=labs.hlthdays),
                                race=factor(race2, labels=labs.race2fmt),
                                rr_atwork=factor(rratwrk2, labels=labs.rr2atwrk),
                                rr_cognitive=factor(rrcognt2, labels=labs.rr2cognt),
                                rr_emptsympt=factor(rremtsm2, labels=labs.rr2yesno),
                                rr_physympt=factor(rrphysm2, labels=labs.rr2yesno),
                                rr_hcare=factor(rrhcare3, labels=labs.rr3hcare),
                                rr_assignedrace=factor(rrclass2, labels=labs.rrclass),
                                sex=factor(sex, labels=labs.sex),
                                smoker=factor(smoker, labels=labs.smoker),
                                state=as.character(`_state`), psu=`_psu`,
                                alc_consum=as.numeric(alcpmo), 
                                wt=as.numeric(`_finalwt`)*(n()/sum(as.numeric(`_finalwt`))),
                                year=as.factor(iyearc), 
                                region=as.character(region),
                                strata=as.factor(`_ststr`)) %>%
      rownames_to_column(var="ID") %>%
      select(ID, age_gr, gen_health, educ, hvy_drink, income, 
             ment_days, phys_days, race, rr_atwork, rr_cognitive, rr_emptsympt,
             rr_physympt, rr_hcare, rr_assignedrace, sex, smoker, year, state, region,
             alc_consum, wt, strata)
```

#Sample Profiles by State and Year
```{r warning=FALSE, message=FALSE}
library(tableone); library(survey)

#gather separate data frames for states as list
dfs <- split(brfss.all, brfss.all$state) 

#define sample profile variables
sample.profile <- c("sex", "age_gr", "educ", "income", "race", "region",
               "gen_health", "ment_days", "phys_days",  
               "hvy_drink", "alc_consum", "smoker", 
               "rr_cognitive", "rr_atwork", "rr_emptsympt", "rr_physympt", "rr_hcare",
               "rr_assignedrace") 

#Unweighted sample profile table
sample.descript <- CreateTableOne(sample.profile, data=brfss.all, test=F)

#weighted sample profile table
#creates yearBYstrata strata variable
brfss.all <- brfss.all %>% mutate(strata_all = interaction(strata,  year, drop = TRUE ))

#sets survey design
brfss_svy <- svydesign(ids = ~ ID, strata = ~ strata_all, weights = ~ wt, data = brfss.all)

#creates table
svysample.descript <- svyCreateTableOne(sample.profile, data=brfss_svy, test=F)
```

###Unweighted
```{r}
sample.descript
```

###Weighted
```{r}
svysample.descript
```

#Initial item and scale analysis
```{r initial item analysis, message=FALSE, warning=FALSE}
#recode rr variables
rr.df <- brfss.all %>% dplyr::select(strata_all, wt, ID, contains("rr"), -rr_assignedrace) %>%
      mutate(
            hcare = fct_collapse(rr_hcare, NULL = c("No health care in past 12 months",
                                                  "Only encountered people of the same race")),
            physympt = fct_collapse(rr_physympt, NULL = c("Don't know/Not Sure", "Refused")), 
            emptsympt = fct_collapse(rr_emptsympt, NULL = c("Don't know/Not Sure", "Refused")),
            at_work = fct_collapse(rr_atwork, NULL = c("Only encountered people of the same race")),
            cognitive = rr_cognitive
      ) %>%
      mutate(hcare=fct_relevel(hcare, "Worse than some races, better than others", after = 1),
             at_work=fct_relevel(at_work, "Worse than some races, better than others", after = 1)) %>%
      dplyr::select(-contains("rr"))  %>%
      mutate_at(vars(hcare, physympt, emptsympt, at_work), fct_rev) %>%
      mutate_at(vars(hcare, physympt, emptsympt, at_work, cognitive), as.ordered) 

#fit models
library(psych)
#scale sampling weight to sample size
rr.dfnum <- rr.df %>% mutate_all(as.numeric)

mod1 <- fa(rr.dfnum[, 4:8], fm="minres", weight=rr.dfnum$wt, cor="poly")

corPlot(mod1$r, upper=F, numbers=T, main="Polychoric correlations, weighted")

c_alpha <- psych::alpha(mod1$r)
```

The figure shows the weighted polychoric correlations. Cognitive does not have large correlations with the remaining variables, so it is likely to not perform so well in subsequent psychometric modeling. We'll keep an eye on it and see if it is important in other ways (e.g. by having a distinct location/severity parameter).

The scale reliability of `r round(c_alpha$total[,2], 3)` is decent, so is the average correlation of `r round(c_alpha$total[,4], 3)`

```{r}
round(c_alpha$total, 2)
```

As you can see below, this is would incrase to `r round(c_alpha$alpha.drop[5,2], 3)` with an average correlation of `r round(c_alpha$alpha.drop[5,4], 3)` if cognitive is dropped.

```{r}
round(c_alpha$alpha.drop, 3)
```

Not surprisingly, cognitive has a small item-total correlation of `r round(c_alpha$item.stats[5,2], 3)`

```{r}
round(c_alpha$item.stats, 3)
```


This general pattern is confirmed in the factor analysis, where cognitive has a smaller factor loading relative to the others.

```{r}
print(mod1)
```


# Some outstanding questions before proceeding:

1. How does "other's perceive your race" fit into this measurement model? A key assumption of measurement models is that the latent construct **causes** the observed responses to the items.  There is a perception component to it, but I struggle to see how socially assigned race fits with the other items in a *common cause* model. I've left it out of the psychometric analysis for now. We should discuss whether it fits in the scale or is part of a structural model.

2. Is "better than..." to "worse than other races" and ordinal or nominal scale? Specfically, can we assign an order to "worse than some races, better than others" vs. "The same as other races".  I'm treating it as ordinal for now, but I can see an argument for it not being ordinal.

3. Treating "have only seen people same race" and "no health care in past 12 months" (and similar) items as missing. I've treated them as missing, but they can be modeled as a conditional category. I prefer the former, but we should discuss.

4. Is the behavior of cognitive interpretable and expected?

5. Modeling goals. We outlined the DIF analyses, but do you want to assess conditional reliablity and inspect "severity/location" parameters?
